\huge Open, privacy-preserving protocols for lawful surveillance

### 2016-02-29 paper/chaining.tex
small edits plus corrected description of Kearns's results

\section{Lawful Contact Chaining}\label{sec-chaining}
Contact chaining is a form of government surveillance with which it is deceptively easy to expose many innocent, untargeted users to government scrutiny. The goal of contact chaining is to use information about social connections between identities, such as records of phone calls between one number and another, to identify members of a criminal organization or terrorist group. Starting with one or more suspects whose identities are known, the government aims to consider contacts of those suspects. These can be \emph{direct contacts}, such as two people who spoke on the phone, or \emph{extended contacts}, such as two people connected by a chain of two or more phone calls. If Alice calls Bob, and Bob calls Charlie, then Alice and Bob are direct contacts (as are Bob and Charlie), but Alice and Charlie are extended contacts. We may also say that Alice and Charlie are at a distance of 2 in the communication graph (since the shortest number of phone calls that connect Alice to Charlie is 2).

@@

%This is a candidate protocol for privacy-preserving contact chaining. We discuss the desired inputs, outputs, and security assumptions below, and then present two versions of the protocol in sections \ref{sec-proto1} and \ref{sec-proto2}.

@@

We specify the protocol in full in sections~\ref{sec-proto1} and \ref{sec-proto2}.

@@

The agencies' outputs from the protocols in sections~\ref{sec-proto1} and \ref{sec-proto2} will be $\mathbf{C}$. $\mathbf{C}$ will contain agency ciphertexts of all phone numbers at most $k$ phone calls away from $x$, considering only vertices of degree at most $d$. This is the desired output. $C$ reveals nothing to any agencies unless they all provide their decryption keys. To continue the process of lawful investigation, the agencies should combine the output $\mathbf{C}$ with other sets of potential suspects (such as from further runs of this protocol, or from cell tower dumps) in a lawful intersection protocol.

>>>>

\section{Lawful Contact Chaining}\label{sec-chaining}
Contact chaining is a form of government surveillance with which it is deceptively easy to expose many innocent, untargeted users to government scrutiny. The goal of contact chaining is to use information about social connections between identities, such as records of phone calls between one number and another, to identify members of a criminal organization or terrorist group. Starting with one or more suspects whose identities are known, the government aims to consider contacts of those suspects. These can be \emph{direct contacts}, such as two people who spoke on the phone, or \emph{extended contacts}, such as two people connected by a chain of two or more phone calls. If Alice calls Bob, and Bob calls Charlie, then Alice and Bob are direct contacts (as are Bob and Charlie), but Alice and Charlie are extended contacts. We may also say that Alice and Charlie are at distance 2 in the communication graph (because the smallest number of phone calls that connect Alice to Charlie is 2).

@@

%This is a candidate protocol for privacy-preserving contact chaining. We discuss the desired inputs, outputs, and security assumptions below, and then present two versions of the protocol in Sections \ref{sec-proto1} and \ref{sec-proto2}.

@@

We specify the protocol in full in Sections~\ref{sec-proto1} and \ref{sec-proto2}.

@@

The agencies' outputs from the protocols in Sections~\ref{sec-proto1} and \ref{sec-proto2} will be $\mathbf{C}$. $\mathbf{C}$ will contain agency ciphertexts of all phone numbers at most $k$ phone calls away from $x$, considering only vertices of degree at most $d$. This is the desired output. $C$ reveals nothing to any agencies unless they all provide their decryption keys. To continue the process of lawful investigation, the agencies should combine the output $\mathbf{C}$ with other sets of potential suspects (such as from further runs of this protocol, or from cell tower dumps) in a lawful intersection protocol.

### 2016-02-29 paper/intro.tex
small edits plus corrected description of Kearns's results

In Section~\ref{sec-chaining}, we turn our attention to 
accountable-surveillance protocols for contact chaining.
The goal of contact chaining is to use the topology of a {\it communication 
graph} ({\it e.g.}, a phone-call graph, email graph, or social network) 
in order to identify associates (or ``contacts'') of lawfully targeted users. 
Government agencies can then investigate those associates to determine whether
they deserve further attention.\footnote{Note that contact chaining is not
tantamount to ``guilt by association.''  Rather, it is tantamount to
``suspicion by association,'' which is in fact a time-honored principal in
law enforcement.  When investigating a murder, rape, or other violent felony,
police departments first turn their attention to the associates of the victim.
Accountable-surveillance practices require that contacts who are suspected,
investigated, and found {\it not} to be involved in illegal activity be 
cleared of suspicion and that their personal information be deleted from 
agencies' databases.} It is useful to consider both direct contacts, 
{\it i.e.}, users who are neighbors in the communication graph, and
extended contacts, {\it i.e.}, users who are at distance $k$ in the 
communication graph, for an appropriate constant $k$.
In a phone-call graph, if Alice calls Bob, 
and Bob calls Charlie, then Alice and Bob are direct contacts (as are Bob and 
Charlie), but Alice and Charlie are extended contacts (more precisely,
contacts at distance 2).  Without accountability and security mechanisms to
limit an investigation's scope, contact chaining in a mass-communication
network can sweep in a huge number of untargeted users.
In Section~\ref{sec-chaining}, we provide an accountable contact-chaining
protocol that bounds the scope of the search, uses encryption to protect 
untargeted users, and is computationally efficient (in that its time
complexity and space complexity are both linear in the size of the output).

>>>>

In Section~\ref{sec-chaining}, we turn our attention to 
accountable-surveillance protocols for contact chaining.
The goal of contact chaining is to use the topology of a {\it communication 
graph} ({\it e.g.}, a phone-call graph, email graph, or social network) 
in order to identify associates (or ``contacts'') of lawfully targeted users. 
Government agencies can then investigate those associates to determine whether
they deserve further attention.\footnote{Note that contact chaining is not
tantamount to ``guilt by association.''  Rather, it is tantamount to
``suspicion by association,'' which is in fact a time-honored principal in
law enforcement.  When investigating a murder, rape, or other violent felony,
police departments first turn their attention to the associates of the victim.
Accountable-surveillance practices require that contacts who are suspected,
investigated, and found {\it not} to be involved in illegal activity be 
cleared of suspicion and that their personal information be deleted from 
agencies' databases.} It is useful to consider both direct contacts, 
{\it i.e.}, users who are neighbors in the communication graph, and
extended contacts, {\it i.e.}, users who are at distance $k$ in the 
communication graph, for an appropriate constant $k$.
In a phone-call graph, if Alice calls Bob, 
and Bob calls Charlie, then Alice and Bob are direct contacts (as are Bob and 
Charlie), but Alice and Charlie are extended contacts (more precisely,
contacts at distance 2).  Without accountability and security mechanisms to
limit an investigation's scope, contact chaining in a mass-communication
network can sweep in a huge number of untargeted users.
In Section~\ref{sec-chaining}, we provide an accountable contact-chaining
protocol that bounds the scope of the search, uses encryption to protect 
untargeted users, and is computationally efficient (in that its time
complexity and communication complexity are both linear in the size of the 
output).

### 2016-02-29 paper/related.tex
small edits plus corrected description of Kearns's results

\section{Related Work}\label{sec-related}
Privacy-preserving computation has been studied extensively. An overview
of the approaches taken by the cryptographic-research community is provided by
Perry {\it et al.}~\cite{pgfw-scn2014}; an overview of of the data-mining
approach is provided by Aggarwal and Yu~\cite{ay2008}.  Kamara~\cite{kamara}
and Kroll {\it et al.}~\cite{kroll} used cryptographic protocols to 
achieve privacy and accountability in the surveillance of known targets.
Segal {\it et al.}~\cite{sff-foci2014} formulated the openness principle that
we have followed and were the first to design privacy-preserving protocols for
the surveillance of unknown targets.  Kearns {\it et al.}~\cite{krwy-pnas16}
present efficient graph-search algorithms that distinguish targeted users from
untargeted users while preserving the privacy of direct contacts of untargeted
users; unlike our privacy-preserving contact-chaining algorithms, which rely
on cryptographic techniques, their search algorithms rely on 
differential-privacy techniques.

>>>>

\section{Related Work}\label{sec-related}
Privacy-preserving computation has been studied extensively. An overview
of the approaches taken by the cryptographic-research community is provided by
Perry {\it et al.}~\cite{pgfw-scn2014}; an overview of of the data-mining
approach is provided by Aggarwal and Yu~\cite{ay2008}.

Kamara~\cite{kamara}
and Kroll {\it et al.}~\cite{kroll} used cryptographic protocols to 
achieve privacy and accountability in the surveillance of known targets.
Segal {\it et al.}~\cite{sff-foci2014} formulated the openness principle that
we have followed and were the first to design privacy-preserving protocols for
the surveillance of unknown targets.

Kearns {\it et al.}~\cite{krwy-pnas16} present efficient graph-search 
algorithms that distinguish targeted users from untargeted users; for each 
untargeted user $u$, the set of direct contacts of $u$ remains private.
Unlike our privacy-preserving contact-chaining algorithms, which rely
on cryptographic techniques, their graph-search algorithms rely on 
differential-privacy techniques.

### 2016-02-29 paper/submission.tex
Fix graph in Sec. 3
Attempt to fix compile error
small edits plus corrected description of Kearns's results

\documentclass[USenglish,oneside,twocolumn]{article}

\usepackage[utf8]{inputenc}%(only for the pdftex engine)
%\RequirePackage[no-math]{fontspec}%(only for the luatex or the xetex engine)
\usepackage[big]{dgruyter_NEW}

@@

\usepackage{subcaption}

@@

\bibitem{verdict} H.~Corrigan-Gibbs, D.~I.~Wolinsky, and B.~Ford, In: 
S.~King (Ed.), $22^{nd}$ Security Symposium, August 14-16, 2013, 
Washington, D.C., USA (USENIX, Berkeley, 2013)

\bibitem{snapnets} J.~Leskovec and A.~Krevl, SNAP Datasets: 
Stanford Large Network Dataset Collection, 
http://snap.stanford.edu/data

@@

\bibitem{pgfw-scn2014} J. Perry, D. Gupta, J. Feigenbaum, and R. Wright,
In: M. Abdalla and R. De Prisco (eds.), $9^{th}$ Conference on Security and 
Cryptography for Networks, September 3-5, 2014, Amalfi, Italy (Springer,
Berlin, 2014)

\bibitem{ay2008} C.~Aggarwal and P.~Yu, In: C.~Aggarwal and P.~Yu (Eds.), 
Privacy-Preserving Data Mining: Models and Algorithms 
(Springer, Berlin, 2008) 11

@@

\bibitem{kroll} J.~Kroll, E. Felten, and D. Boneh, Secure protocols for
accountable warrant execution, April 2, 2014,\\
https://freedom-to-tinker.com/blog/felten/\\
secure-protocols-for-accountable-warrant-execution/

>>>>

\documentclass[USenglish,oneside,twocolumn]{article}
\usepackage{etex}
\usepackage[utf8]{inputenc}%(only for the pdftex engine)
%\RequirePackage[no-math]{fontspec}%(only for the luatex or the xetex engine)
\usepackage[big]{dgruyter_NEW}

@@

\usepackage{subcaption}
\usepackage{pgfplots}

@@

\bibitem{verdict} H.~Corrigan-Gibbs, D.~I.~Wolinsky, B.~Ford, In: 
S.~King (Ed.), $22^{nd}$ Security Symposium, August 14-16, 2013, 
Washington, D.C., USA (USENIX, Berkeley, 2013)

\bibitem{snapnets} J.~Leskovec, A.~Krevl, SNAP Datasets: 
Stanford Large Network Dataset Collection, 
http://snap.stanford.edu/data

@@

\bibitem{pgfw-scn2014} J. Perry, D. Gupta, J. Feigenbaum, R. Wright,
In: M. Abdalla and R. De Prisco (eds.), $9^{th}$ Conference on Security and 
Cryptography for Networks, September 3-5, 2014, Amalfi, Italy (Springer,
Berlin, 2014)

\bibitem{ay2008} C.~Aggarwal, P.~Yu, In: C.~Aggarwal and P.~Yu (Eds.), 
Privacy-Preserving Data Mining: Models and Algorithms 
(Springer, Berlin, 2008) 11

@@

\bibitem{kroll} J.~Kroll, E. Felten, D. Boneh, Secure protocols for
accountable warrant execution, April 2, 2014,\\
https://freedom-to-tinker.com/blog/felten/\\
secure-protocols-for-accountable-warrant-execution/
